Stack Safety is a Security Property

### 2021-04-17 tex/main.tex
Fixed passing predicate

10 f:   add $2,%sp      ; allocate frame
11      mov $5,-2(%sp)  ; assign into y
12      mov -4(%sp),-1(%sp)  ; violation!
13      mov -2(%sp),%ra ; calculate and
14      add -1(%sp),%ra ; set return value
15      sub $2,%sp      ; deallocate frame
16      ret
\end{verbatim}
}
\end{subfigure}
\begin{subfigure}{.59\textwidth}
\centering
\begin{tabular}{l l | l}
  {\tt 2} &
  \multicolumn{2}{c}{
    \memoryaddrs{8em}
    \memory{4}{\unsealc}
    ~$\cdots$
%    \MemoryLabel{-18em}{0.75em}{0}
%    \MemoryLabel{-14em}{0.75em}{0}
%    \MemoryLabel{-10em}{0.75em}{0}
%    \MemoryLabel{-7em}{0.75em}{0}
    \vspace{.5em}
  } \\
  {\tt 11} &
  \memoryaddrs{21em}
  \memory{1}{\unsealc}
  \memory{1}{\retptrc}
  \memory{2}{\unsealc}
  ~$\cdots$
  \MemoryLabel{-19em}{0.75em}{42}
%  \MemoryLabel{-10em}{0.75em}{0}
%  \MemoryLabel{-6em}{0.75em}{0}
  &
  \memoryaddrs{21em}
  \memory{1}{\unsealc}
  \memory{1}{\retptrc}
  \memory{2}{\unsealc}
  ~$\cdots$
  \MemoryLabel{-19em}{0.75em}{\(v_0\)}
  \MemoryLabel{-10em}{0.75em}{\(v_1\)}
  \MemoryLabel{-6em}{0.75em}{\(v_2\)}
  \\
  {\tt 12} &
  \memoryaddrs{21em}
  \memory{1}{\unsealc}
  \memory{1}{\retptrc}
  \memory{1}{\goodc}
  \memory{1}{\unsealc}
  ~$\cdots$
  \MemoryLabel{-19em}{0.75em}{42}
  \MemoryLabel{-10em}{0.75em}{5}
 % \MemoryLabel{-6em}{0.75em}{0}
  &
  \memoryaddrs{21em}
  \memory{1}{\unsealc}
  \memory{1}{\retptrc}
  \memory{1}{\goodc}
  \memory{1}{\unsealc}
  ~$\cdots$
  \MemoryLabel{-19em}{0.75em}{\(v_0\)}
  \MemoryLabel{-10em}{0.75em}{5}
  \MemoryLabel{-6em}{0.75em}{\(v_2\)}
  \\
  {\tt 13} &
  \memoryaddrs{21em}
  \memory{1}{\unsealc}
  \memory{1}{\retptrc}
  \memory{1}{\unsealc}
  \memory{1}{\badc}
  ~$\cdots$
  \MemoryLabel{-19em}{0.75em}{42}
  \MemoryLabel{-10em}{0.75em}{5}
  \MemoryLabel{-6em}{0.75em}{42}
  \vspace{.5em}
  &
  \memoryaddrs{21em}
  \memory{1}{\unsealc}
  \memory{1}{\retptrc}
  \memory{1}{\unsealc}
  \memory{1}{\badc}
  ~$\cdots$
  \MemoryLabel{-19em}{0.75em}{\(v_0\)}
  \MemoryLabel{-10em}{0.75em}{5}
  \MemoryLabel{-6em}{0.75em}{\(v_0\)}
  \vspace{.5em}
\end{tabular}

@@

\paragraph{Formalization of variable passing.}
We add extra information to our call
annotations to reflect that they may now pass variables on the stack, and that the locations
of those variables may be dynamic (as in the case of pass-by-reference).
A call annotation now includes a predicate relating machine states to the components
that should be passed, expressed as offsets from the stack pointer. \apt{Much better, but the
actual definition of $P$ is still wrong.}

\[\codemap \in \CODEMAPS ::= \WORDS \rightharpoonup \callmap(P) \hspace*{1in} P \subseteq (\MACHS \times \COMPONENTS)\]
We extend the set of domains to include passed components,
which are labeled with the range of depths of the caller that has passed them.
\[\stackDom \in \STKDOMS ::= \outside + \unsealed + \sealed{\depth} + \passed{\depth,\depth'}\]
The definition of context looks the same, but is based on the revised definition of \(\STKDOMS\):
\[(\domMap,\rts) = \context \in \CONTEXTS ::= (\COMPONENTS \rightarrow \STKDOMS)
\times (\listT (\target)) \]
The initial context again maps all components to \(\unsealed\).
We straightforwardly describe how the sealing convention and passing predicate combine to
update a domain map at a call from depth \(\depth\).
\[(\mathit{push} ~ \domMap ~ \sealcon ~ P ~ \mach ~ \depth) ~ \component =
\begin{cases}
  \sealed{\depth}              & \text{if } \domMap ~ \component = \unsealed \text{ and }
                                 \sealcon ~ \mach ~ \component \text{ and } \neg P ~ \mach ~ \component \\
  \passed{\depth,\depth+1}     & \text{if } \domMap ~ \component = \unsealed \text{ and }
                                 \sealcon ~ \mach ~ \component \text{ and } P ~ \mach ~ \component \\
  \passed{\depth',\depth+1}    & \text{if } \domMap ~ \component = \passed{\depth',\depth} \text{ and }
                                 \sealcon ~ \mach ~ \component \text{ and } P ~ \mach ~ \component,
                                 \text{ for any } \depth' \\
  \domMap~\component & \text{else} \\
\end{cases}\]

At the matching return we just unseal the components at depth \(\depth\):

\[(\mathit{pop} ~ \domMap ~ \depth) ~ \component =
\begin{cases}
  \unsealed                 & \text{if } \domMap ~ \component = \sealed{\depth} \\
  \unsealed                 & \text{if } \domMap ~ \component = \passed{\depth-1,\depth} \\
  \passed{\depth',\depth-1} & \text{if } \domMap ~ \component = \passed{\depth',\depth}
                              \text{ for any } \depth' < \depth \\
  \domMap~\component        & \text{else} \\
\end{cases}\]

@@

We have introduced a characterization against which stack safety enforcement
mechanisms can be measured and used it to evaluate tag-based enforcement mechanisms.
From here there are two natural ways to extend the work. We should thoroughly evaluate
other mechanisms, especially Cheri-based capability systems. And we can further demonstrate
the functionality of our definitions by showing that they can be used to prove
the correctness of a programming logic with strengthened rules around calls and returns.

>>>>

10 f:   add $2,%sp      ; allocate frame
11      mov $5,-2(%sp)  ; assign into y
12      mov -4(%sp),-1(%sp)  ; violation!
13      mov -2(%sp),%ra ; calculate and
14      add -1(%sp),%ra ; set return value
15      sub $2,%sp      ; deallocate frame
16      ret
\end{verbatim}
}
\end{subfigure}
\begin{subfigure}{.59\textwidth}
\centering
\begin{tabular}{l l | l}
  {\tt 2} &
  \multicolumn{2}{c}{
    \memoryaddrs{8em}
    \memory{4}{\unsealc}
    ~$\cdots$
%    \MemoryLabel{-18em}{0.75em}{0}
%    \MemoryLabel{-14em}{0.75em}{0}
%    \MemoryLabel{-10em}{0.75em}{0}
%    \MemoryLabel{-7em}{0.75em}{0}
    \vspace{.5em}
  } \\
  {\tt 11} &
  \memoryaddrs{21em}
  \memory{1}{\unsealc}
  \memory{1}{\retptrc}
  \memory{2}{\unsealc}
  ~$\cdots$
  \MemoryLabel{-19em}{0.75em}{42}
%  \MemoryLabel{-10em}{0.75em}{0}
%  \MemoryLabel{-6em}{0.75em}{0}
  &
  \memoryaddrs{21em}
  \memory{1}{\unsealc}
  \memory{1}{\retptrc}
  \memory{2}{\unsealc}
  ~$\cdots$
  \MemoryLabel{-19em}{0.75em}{\(v_0\)}
  \MemoryLabel{-10em}{0.75em}{\(v_1\)}
  \MemoryLabel{-6em}{0.75em}{\(v_2\)}
  \\
  {\tt 12} &
  \memoryaddrs{21em}
  \memory{1}{\unsealc}
  \memory{1}{\retptrc}
  \memory{2}{\unsealc}
  ~$\cdots$
  \MemoryLabel{-19em}{0.75em}{42}
  \MemoryLabel{-10em}{0.75em}{5}
 % \MemoryLabel{-6em}{0.75em}{0}
  &
  \memoryaddrs{21em}
  \memory{1}{\unsealc}
  \memory{1}{\retptrc}
  \memory{2}{\unsealc}
  ~$\cdots$
  \MemoryLabel{-19em}{0.75em}{\(v_0\)}
  \MemoryLabel{-10em}{0.75em}{5}
  \MemoryLabel{-6em}{0.75em}{\(v_2\)}
  \\
  {\tt 13} &
  \memoryaddrs{21em}
  \memory{1}{\unsealc}
  \memory{1}{\retptrc}
  \memory{1}{\unsealc}
  \memory{1}{\badc}
  ~$\cdots$
  \MemoryLabel{-19em}{0.75em}{42}
  \MemoryLabel{-10em}{0.75em}{5}
  \MemoryLabel{-6em}{0.75em}{42}
  \vspace{.5em}
  &
  \memoryaddrs{21em}
  \memory{1}{\unsealc}
  \memory{1}{\retptrc}
  \memory{1}{\unsealc}
  \memory{1}{\badc}
  ~$\cdots$
  \MemoryLabel{-19em}{0.75em}{\(v_0\)}
  \MemoryLabel{-10em}{0.75em}{5}
  \MemoryLabel{-6em}{0.75em}{\(v_0\)}
  \vspace{.5em}
\end{tabular}

@@

\paragraph{Formalization of variable passing.}
We add extra information to our call
annotations to reflect that they may now pass variables on the stack, and that the locations
of those variables may be dynamic (as in the case of pass-by-reference).
A call annotation now includes a predicate relating machine states to the components
that should be passed, expressed as offsets from the stack pointer.
%
\[\codemap \in \CODEMAPS ::= \WORDS \rightharpoonup \callmap(P \subseteq \WORDS)\]

We extend the set of domains to include passed components,
which are labeled with the range of depths of the caller that has passed them.
\[\stackDom \in \STKDOMS ::= \outside + \unsealed + \sealed{\depth} + \passed{\depth,\depth'}\]
The definition of context looks the same, but is based on the revised definition of \(\STKDOMS\):
\[(\domMap,\rts) = \context \in \CONTEXTS ::= (\COMPONENTS \rightarrow \STKDOMS)
\times (\listT (\target)) \]
The initial context again maps all components to \(\unsealed\).
We straightforwardly describe how the sealing convention and passing predicate combine to
update a domain map at a call from depth \(\depth\).
\[(\mathit{push} ~ \domMap ~ \sealcon ~ P ~ \mach ~ \depth) ~ \component =
\begin{cases}
  \sealed{\depth}              & \text{if } \domMap ~ \component = \unsealed \text{ and }
                                 \sealcon ~ \mach ~ \component \text{ and }
                                 \component - (\mach ~ \rsp) \not \in P \\
  \passed{\depth,\depth+1}     & \text{if } \domMap ~ \component = \unsealed \text{ and }
                                 \sealcon ~ \mach ~ \component \text{ and }
                                 \component - (\mach ~ \rsp) \in P \\
  \passed{\depth',\depth+1}    & \text{if } \domMap ~ \component = \passed{\depth',\depth} \text{ and }
                                 \sealcon ~ \mach ~ \component \\
                               & \text{ and } (\mach ~ \rsp) \in P, \text{ for any } \depth' \\
  \domMap~\component & \text{else} \\
\end{cases}\]

At the matching return we just unseal the components at depth \(\depth\):
%
\[(\mathit{pop} ~ \domMap ~ \depth) ~ \component =
\begin{cases}
  \unsealed                 & \text{if } \domMap ~ \component = \sealed{\depth} \\
  \unsealed                 & \text{if } \domMap ~ \component = \passed{\depth-1,\depth} \\
  \passed{\depth',\depth-1} & \text{if } \domMap ~ \component = \passed{\depth',\depth}
                              \text{ for any } \depth' < \depth \\
  \domMap~\component        & \text{else} \\
\end{cases}\]

@@

We have introduced a characterization against which stack safety enforcement
mechanisms can be measured and used it to evaluate tag-based enforcement mechanisms.
Our immediate next step is to implement testing of observational properties,
both to validate them and to explore how to effectively test properties that do not
fail immediately.
From there there are two natural ways to extend the work. We should thoroughly evaluate
other mechanisms, especially Cheri-based capability systems. And we can further demonstrate
the functionality of our definitions by showing that they can be used to prove
the correctness of a programming logic with strengthened rules around calls and returns.
